---
title: "CSIT165_Project2"
author: "Alex Yu and Audrey Lewis"
date: "12/19/2020"
output: html_document
urlcolor: blue
---

```{r Library, include=FALSE}
#for wget
library(HelpersMG)
#for reading in data
library(readr)
#install stringr
library(stringr)
#install dplyr
library(dplyr)
#install tidyr
library(tidyr)
#install magrittr
library(magrittr)
#install data table
library(data.table)
#install Kable
library(kableExtra)
#install for visualization
library(ggplot2)
#install for date function
library(lubridate)
#install tidyverse
library(tidyverse)
#install cowplot
library(cowplot)

knitr::opts_chunk$set(echo = TRUE)
```


```{r Files, include=FALSE}

#Reading files for confirmedGlobal
# urlfile="./time_series_covid19_confirmed_global.csv"
# ConfirmedGlobal <- read_csv(url(urlfile))
ConfirmedGlobal <- read_csv("./time_series_covid19_confirmed_global.csv")

#Reading files for DeathsGlobal
#urlfile="./time_series_covid19_deaths_global.csv"
#DeathsGlobal <- read_csv(url(urlfile))
DeathsGlobal <- read_csv("./time_series_covid19_deaths_global.csv")

#Reading files for RecoveredGlobal
# urlfile="./time_series_covid19_recovered_global.csv"
# RecoveredGlobal <- read_csv(url(urlfile))
RecoveredGlobal <- read_csv("./time_series_covid19_recovered_global.csv")

#Reading files for confirmedGlobal
# urlfile="./time_series_covid19_confirmed_US.csv"
# ConfirmedUS <- read_csv(url(urlfile))
ConfirmedUS <- read_csv("./time_series_covid19_confirmed_US.csv")

#Reading files for DeathsGlobal
# urlfile="./time_series_covid19_deaths_US.csv"
# DeathsUS <- read_csv(url(urlfile))
DeathsUS <- read_csv("./time_series_covid19_deaths_US.csv")

```


<!-- testing section only -->
------

# Testing Section Only

```{r testUpload-1, echo=FALSE}
head(ConfirmedGlobal)
```

If the ConfirmedGlobal table was loaded correctly, you should see a tibble above with that data.

```{r testUpload-2, echo=FALSE}
head(DeathsGlobal)
```

If the DeathsGlobal table was loaded correctly, you should see a tibble above with that data.

```{r testUpload-3, echo=FALSE}
head(RecoveredGlobal)
```

If the RecoveredGlobal table was loaded correctly, you should see a tibble above with that data.

```{r testUpload-4, echo=FALSE}
head(ConfirmedUS)
```

If the ConfirmedUS table was loaded correctly, you should see a tibble above with that data.

```{r testUpload-5, echo=FALSE}
head(DeathsUS)
```

If the DeathsUS table was loaded correctly, you should see a tibble above with that data.

------
<!-- end testing section -->



# Objective 1 - Global Trajectory

<!-- assigned to Audrey -->
<!-- enter code here -->


# Objective 2 - Global Map

<!-- assigned to Audrey -->
<!-- enter code here -->


# Objective 3 - Narrowing Down Hot Spots

<!-- assigned to Audrey -->
<!-- enter code here -->


# Objective 4 - Zooming Into California

<!-- Chart 1 -->

```{r obj4-code1, echo=TRUE}
# creates new data frame for only CA confirmed cases
dfConfirmedCAChartData <- ConfirmedUS %>%
  
  # removes all columns with numeric values that aren't case counts
  select(-(`UID`)) %>%
  select(-(`code3`)) %>%
  select(-(`FIPS`)) %>%
  select(-(`Lat`)) %>%
  select(-(`Long_`)) %>%
  
  # retains only the California row
  filter(`Province_State` == "California") %>%
  
  # groups by Country/Region
  group_by(`Province_State`) %>%
  
  # totals all columns that are numeric
  summarise_if(is.numeric, sum) %>%

  # gathers data into long form
  gather(key=Date, value=`Confirmed Cases`, -`Province_State`) %>%
  
  # converts the date column into real dates for the chart
  mutate(`Date`=as.Date(`Date`, format = "%m/%d/%y"))
```

```{r obj4-code2, echo=TRUE}
# creates new data frame for only CA deaths
dfDeathsCAChartData <- DeathsUS %>%
  
  # removes all columns with numeric values that aren't case counts
  select(-(`UID`)) %>%
  select(-(`code3`)) %>%
  select(-(`FIPS`)) %>%
  select(-(`Lat`)) %>%
  select(-(`Long_`)) %>%
  select(-(`Population`)) %>%
  
  # retains only the California row
  filter(`Province_State` == "California") %>%
  
  # groups by Country/Region
  group_by(`Province_State`) %>%
  
  # totals all columns that are numeric
  summarise_if(is.numeric, sum) %>%
  
  # gathers data into long form
  gather(key=Date, value=`Deaths`, -`Province_State`) %>%
  
  # converts the date column into real dates for the chart
  mutate(`Date`=as.Date(`Date`, format = "%m/%d/%y"))
```

Below is a chart showing the deaths and confirmed cases for the state of California over time. The black vertical line provides a visual representation of when the first shutdowns occurred on March 19, 2020.

```{r obj4-chart1, echo=TRUE}
obj4chart1colors <- c("Confirmed Cases" = "blue", "Deaths" = "red")

ggplot() +
  geom_line(data = dfConfirmedCAChartData,
            aes(x = `Date`, y = `Confirmed Cases`,
            color = "Confirmed Cases")) +
  geom_line(data = dfDeathsCAChartData,
            aes(x = `Date`, y = `Deaths`,
            color = "Deaths")) +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Date", y = "Cases", color = "California Total") +
  ggtitle("California's Trajectory for COVID-19") +
  scale_color_manual(values = obj4chart1colors) +
  geom_vline(xintercept = as.numeric(as.Date("2020-03-19")),
             color = "black")
```


<!-- Chart 2 -->

```{r obj4-code3, echo=TRUE}
# reuses earlier data frame for only CA confirmed cases by county
dfConfirmedCAChartData <- ConfirmedUS %>%
  
  # removes all columns with numeric values that aren't case counts
  select(-(`UID`)) %>%
  select(-(`code3`)) %>%
  select(-(`FIPS`)) %>%
  select(-(`Lat`)) %>%
  select(-(`Long_`)) %>%
  
  # retains only the California rows
  filter(`Province_State` == "California") %>%
  
  # retains only the rows for San Diego, Riverside, LA
  filter(`Admin2` %in%
           c("San Diego", "Riverside", "Los Angeles")) %>%
  
  # groups by County (aka Admin2)
  group_by(`Admin2`) %>%
  
  # totals all columns that are numeric
  summarise_if(is.numeric, sum) %>%

  # gathers data into long form
  gather(key=Date, value=`Confirmed Cases`, -`Admin2`) %>%
  
  # converts the date column into real dates for the chart
  mutate(`Date`=as.Date(`Date`, format = "%m/%d/%y"))
```

Below is a chart showing the confirmed cases over time, only for the counties of Los Angeles, Riverside, and San Diego. The black vertical line provides a visual representation of when the first shutdowns occurred on March 19, 2020.

```{r obj4-chart2, echo=TRUE}
ggplot(dfConfirmedCAChartData,
       aes(x = `Date`, y = `Confirmed Cases`,
           group = `Admin2`)) +
  geom_line(aes(color=`Admin2`)) +
  ggtitle("Southern California Confirmed Cases of COVID-19") +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Date", y = "Cases", color = "Counties") +
  geom_vline(xintercept = as.numeric(as.Date("2020-03-19")),
             color = "black")
```

# Objective 5 - Digging Deeper







<!-- commenting this out until submission
     since it throws an error -->

<!-- ```{bash }
# commenting this out until submission
# since it throws an error
#
# git log --pretty=format:"%nSubject: %s%nAuthor: %aN%nDate: %aD%nBody: %b"

``` -->
